---
title: "Thesis Nur Okta Milatina TISE 2023"
format: html
editor: visual
---

# Packages

```{r}
install.packages('dplyr')
install.packages('ggplot2')
install.packages('plm')
install.packages('lmtest')
install.packages('readxl')
install.packages('openxlsx')
install.packages('tseries')
install.packages('lfe')
install.packages('vars')
install.packages('psych')
install.packages('gridExtra')
install.packages('car')
install.packages('broom')
install.packages('stargazer')
install.packages('olsrr')

```

# Dataset

```{r}
library(readxl)
dataset<-read_xlsx('19_05_2.xlsx',sheet=1)
label<-read_xlsx('19_05_2.xlsx',sheet=2)
def<-read_xlsx('19_05_2.xlsx',sheet=3)
```

```{r}
head(dataset)
tail(dataset)
```

```{r}
library(psych)
data_vars <- dataset[,c(4:19)]

# create summary statistics for each variable
SS <- describe(data_vars)
#reference for kurtosis and skewness 
#https://www.sciencedirect.com/topics/neuroscience/kurtosis

# print summary statistics
print(SS)

# count missing data

write.csv(SS, file = "Data_statistics.csv")
```

```{r}
#Overall pattern of CO2
library(ggplot2)
library(dplyr)
# Define a vector of colors for the countries
country_colors <- rainbow(38)

# Define a vector of shapes for the countries
num_shapes <- 5
num_colors <- length(country_colors)
country_shapes <- rep(0:(num_shapes-1), length.out=num_colors)


plot<-ggplot(data = dataset %>% filter(Year>=1990& Year<=2019), aes(x =Year, y = ln_CO2)) +
  geom_point(aes(color = factor(Country), shape = factor(Country)))+
  geom_line(aes(colour = as.factor(Country))) +
  geom_smooth(method = "lm", se = F, lty = "dashed") +
  scale_color_manual(values = country_colors) +
  scale_shape_manual(values = country_shapes) +
  labs(x = "Year",  y = "log CO2 emission (metric ton per capita)",color = "Country",
       shape = "Country") +
  theme_classic() 


plot + scale_x_continuous(limits = c(1990, 2019),breaks = seq(1990,2019, by = 5))

ggsave("log_CO2.png", plot = plot, width = 8, height = 6, dpi = 300)
```

```{r}
#correlation
# assume you have a correlation matrix called 'cor2'
cor <- round(cor(dataset[, c(4:19)], use = "pairwise.complete.obs"),3)

# write the correlation matrix to a CSV file
write.csv(cor, file = "correlation_matrix.csv")

#from the result it seems like Urban Population does not have strong correlation with the Co2 emission

print(cor)
```

# First test of multicollinearity

This section is part of assumption test to find the right variables for the model

```{r}
library("olsrr")
fe_time_lm <- lm(ln_CO2 ~ ln_IU+ln_MIS+ln_OP+ln_GDP+ln_TR+ln_FDI+ln_LS+ln_LE+ln_ME+ factor(Country) + factor(Year), data=dataset)
ols_vif_tol(fe_time_lm)
ols_eigen_cindex(fe_time_lm)
summary(fe_time_lm)
```

```{r}
cor_2 <- round(cor(dataset[, c(4:28)], use = "pairwise.complete.obs"),3)
print(cor_2)
```

# Selecting variables

```{r}
#FDI
te1<- lm(ln_CO2 ~ ln_IU+ln_FDI+ln_LS+ factor(Country) + factor(Year), data=dataset)
te2<- lm(ln_CO2 ~ ln_IU+ln_FDI+ln_LE+ factor(Country) + factor(Year), data=dataset)
te3<- lm(ln_CO2 ~ ln_IU+ln_FDI+ln_ME+ factor(Country) + factor(Year), data=dataset)
te4<- lm(ln_CO2 ~ ln_MIS+ln_FDI+ln_LS+ factor(Country) + factor(Year), data=dataset)
te5<- lm(ln_CO2 ~ ln_MIS+ln_FDI+ln_LE+ factor(Country) + factor(Year), data=dataset)
te6<- lm(ln_CO2 ~ ln_MIS+ln_FDI+ln_ME+ factor(Country) + factor(Year), data=dataset)
te7<- lm(ln_CO2 ~ ln_OP+ln_FDI+ln_LS+ factor(Country) + factor(Year), data=dataset)
te8<- lm(ln_CO2 ~ ln_OP+ln_FDI+ln_LE+ factor(Country) + factor(Year), data=dataset)
te9<- lm(ln_CO2 ~ ln_OP+ln_FDI+ln_ME+ factor(Country) + factor(Year), data=dataset)
ols_vif_tol(te_1)
ols_vif_tol(te_2)
ols_vif_tol(te_3)
ols_vif_tol(te_4)
ols_vif_tol(te_5)
ols_vif_tol(te_6)
ols_vif_tol(te_7)
ols_vif_tol(te_8)
ols_vif_tol(te_9)
```

```{r}
#GDP
te_1<- lm(ln_CO2 ~ ln_OP+ln_GDP+ln_LE+ factor(Country) + factor(Year), data=dataset)
te_2<- lm(ln_CO2 ~ ln_OP+ln_GDP+ln_LS+ factor(Country) + factor(Year), data=dataset)
te_3<- lm(ln_CO2 ~ ln_OP+ln_GDP+ln_ME+ factor(Country) + factor(Year), data=dataset)
te_4<- lm(ln_CO2 ~ ln_IU+ln_GDP+ln_LE+ factor(Country) + factor(Year), data=dataset)
te_5<- lm(ln_CO2 ~ ln_IU+ln_GDP+ln_LS+ factor(Country) + factor(Year), data=dataset)
te_6<- lm(ln_CO2 ~ ln_IU+ln_GDP+ln_ME+ factor(Country) + factor(Year), data=dataset)
te_7<- lm(ln_CO2 ~ ln_MIS+ln_GDP+ln_LE+ factor(Country) + factor(Year), data=dataset)
te_8<- lm(ln_CO2 ~ ln_MIS+ln_GDP+ln_LS+ factor(Country) + factor(Year), data=dataset)
te_9<- lm(ln_CO2 ~ ln_MIS+ln_GDP+ln_ME+ factor(Country) + factor(Year), data=dataset)
ols_vif_tol(te_1)
ols_vif_tol(te_2)
ols_vif_tol(te_3)
ols_vif_tol(te_4)
ols_vif_tol(te_5)
ols_vif_tol(te_6)
ols_vif_tol(te_7)
ols_vif_tol(te_8)
ols_vif_tol(te_9)

```

```{r}
#Trade openness
te_1<- lm(ln_CO2 ~ ln_OP+ln_TR+ln_LE+ factor(Country) + factor(Year), data=dataset)
te_2<- lm(ln_CO2 ~ ln_OP+ln_TR+ln_LS+ factor(Country) + factor(Year), data=dataset)
te_3<- lm(ln_CO2 ~ ln_OP+ln_TR+ln_ME+ factor(Country) + factor(Year), data=dataset)
te_4<- lm(ln_CO2 ~ ln_IU+ln_TR+ln_LE+ factor(Country) + factor(Year), data=dataset)
te_5<- lm(ln_CO2 ~ ln_IU+ln_TR+ln_LS+ factor(Country) + factor(Year), data=dataset)
te_6<- lm(ln_CO2 ~ ln_IU+ln_TR+ln_ME+ factor(Country) + factor(Year), data=dataset)
te_7<- lm(ln_CO2 ~ ln_MIS+ln_TR+ln_LE+ factor(Country) + factor(Year), data=dataset)
te_8<- lm(ln_CO2 ~ ln_MIS+ln_TR+ln_LS+ factor(Country) + factor(Year), data=dataset)
te_9<- lm(ln_CO2 ~ ln_MIS+ln_TR+ln_ME+ factor(Country) + factor(Year), data=dataset)
ols_vif_tol(te_1)
ols_vif_tol(te_2)
ols_vif_tol(te_3)
ols_vif_tol(te_4)
ols_vif_tol(te_5)
ols_vif_tol(te_6)
ols_vif_tol(te_7)
ols_vif_tol(te_8)
ols_vif_tol(te_9)

```

# Panel data analysis

## Pooled OLS

```{r}
#OLS via PLM
library(plm)
ols_a <- plm(ln_CO2 ~ ln_IU+ln_FDI + ln_LS,data = dataset4, 
                      index = c("Country", "Year"), 
                      effect = "individual", model = "pooling")
summary(ols_a)

```

```{r}
library(plm)
ols_b <- plm(ln_CO2 ~ ln_MIS+ln_FDI + ln_LS,data = dataset4, 
                      index = c("Country", "Year"), 
                      effect = "individual", model = "pooling")
summary(ols_b)

```

## Fixed Effects

```{r}
fixed_a <- plm(ln_CO2 ~ ln_IU+ln_FDI+ln_LS, data=dataset4, model="within")
summary(fixed_a)
fixed_b <- plm(ln_CO2 ~ ln_MIS+ln_FDI+ln_LS, data=dataset4, model="within")
summary(fixed_b)
```

```{r}
fixed_a1 <- plm(ln_CO2 ~ ln_IU+ln_FDI, data=dataset4, model="within")

var_decoupling <- var(fixed_a1$residuals)

# Print the variance
print(var_decoupling)
```

### Time-fixed effects

```{r}
fe_time_a <- plm(ln_CO2 ~ ln_IU+ln_FDI+ln_LS, data = dataset4, 
                   index = c("Country", "Year"), effect = "twoways",
                   model = "within")

fe_time_b<- plm(ln_CO2 ~ ln_MIS+ln_FDI+ln_LS, data = dataset4, 
                   index = c("Country", "Year"), effect = "twoways",
                   model = "within")
summary(fe_time_a)
summary(fe_time_b)
```

### Test TFE and FE

```{r}
#Within model
pFtest(fe_time_a, fixed_a) 

```

```{r}
#Within model
pFtest(fe_time_b, fixed_b) 

```

### Test FE and OLS

```{r}
pFtest(fixed_a, ols_a)
```

```{r}
pFtest(fixed_b, ols_b)
```

## Random Effects

```{r}
random_a <- plm(ln_CO2 ~ ln_IU+ln_FDI+ln_LS, data=dataset4, model="random")
summary(random_a)
random_b <- plm(ln_CO2 ~ ln_MIS+ln_FDI+ln_LS, data=dataset4, model="random")
summary(random_b)
```

```{r}
fe_time_lm_a <- lm(ln_CO2 ~ ln_IU+ln_FDI+ln_LS+ factor(Country) + factor(Year), data=dataset4)

summary(fe_time_lm_a)
fe_time_lm_b <- lm(ln_CO2 ~ ln_MIS+ln_FDI+ln_LS+ factor(Country) + factor(Year), data=dataset4)
summary(fe_time_lm_b)
```

### Test RE and OLS

```{r}
plmtest(ols_a, effect = "individual", type = c("bp"))
```

```{r}
plmtest(ols_b, effect = "individual", type = c("bp"))
```

### Test FE and RE

```{r}
phtest(fixed_a, random_a)
```

```{r}
phtest(fixed_b, random_b)
```

### Test BP-test

```{r}
lmtest::bptest(ln_CO2 ~ ln_IU + ln_FDI + ln_LS + factor(Country), 
               studentize = F, data = dataset4)
```

```{r}
lmtest::bptest(ln_CO2 ~ ln_MIS + ln_FDI + ln_LS + factor(Country), 
               studentize = F, data = dataset4)
```

# Assumption test

## Serial autocorrelation test

```{r}
pbgtest(fe_time_a)
```

```{r}
pbgtest(fe_time_a)
```

## Clustered Robust

```{r}
fe_model_felma <- lfe::felm(ln_CO2 ~ ln_IU + ln_FDI + ln_LS  | Country | 0 | Country, data = dataset4)

summary(fe_model_felma)
```

```{r}
fe_model_felmb <- lfe::felm(ln_CO2 ~ ln_MIS + ln_FDI + ln_LS  | Country | 0 | Country, data = dataset4)

summary(fe_model_felmb)
```

```{r}
# Cluster robust standard error
summary(fe_time_a, 
         vcov = vcovHC(fe_time_a,
                       type = "sss",
                      cluster = "group"))
```

```{r}
# Cluster robust standard error
summary(fe_time_b, 
         vcov = vcovHC(fe_time_b,
                       type = "sss",
                      cluster = "group"))
```

## Unit root test

```{r}
pdata<-pdata.frame(dataset,index=c("Country","Year"))

```

```{r}
#define data as panel data
panel <- pdata.frame(dataset4,index("Country","Year"))

head(panel)
```

```{r}
CO2<-purtest(ln_FDI~trend,data=pdata,lags="AIC",pmax=2,test="ips")
summary(CO2)
print(CO2)
```

```{r}
LS_f<-purtest(ln_FDI~trend,data=pdata,lags="AIC",pmax=2,test="madwu")
summary(LS_f)
print(LS_f)
```

```{r}
FDI<-purtest(ln_FDI~trend,data=pdata,lags="AIC",pmax=2,test="Pm")
summary(FDI)
print(FDI)
```

# 
